The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-03-08 src/ms.tex
Add text for Astropy Roadmap section (#22)
* astropy roadmap

* Apply suggestions from code review

Co-authored-by: Adrian Price-Whelan <adrianmpw@gmail.com>

% Links to documentation: subpackages
\newcommand{\astropysubpkg}[1]{\href{http://docs.astropy.org/en/stable/#1/index.html}{\texttt{astropy.#1}}\xspace}
\newcommand{\astropyiosubpkg}[1]{\href{http://docs.astropy.org/en/stable/io/#1/index.html}{\texttt{astropy.io.#1}}\xspace}
\newcommand{\astropywcsaxes}{\href{http://docs.astropy.org/en/stable/visualization/wcsaxes/index.html}{\texttt{astropy.visualization.wcsaxes}}\xspace}
\newcommand{\astropycosmology}{\astropysubpkg{cosmology}}
\newcommand{\astropycosmologyunits}{\href{https://docs.astropy.org/en/stable/cosmology/units.html}{\texttt{astropy.cosmology.units}}}
\newcommand{\astropyunits}{\astropysubpkg{units}}
\newcommand{\astropycoordinates}{\astropysubpkg{coordinates}}
\newcommand{\astropyconstants}{\astropysubpkg{constants}}
\newcommand{\astropytable}{\astropysubpkg{table}}
\newcommand{\astropytime}{\astropysubpkg{time}}
\newcommand{\astropymodeling}{\astropysubpkg{modeling}}
\newcommand{\astropywcs}{\astropysubpkg{wcs}}
\newcommand{\astropyfits}{\astropyiosubpkg{fits}}

@@

\newcommand{\astropySpectralCoord}{\astropyapidoc{coordinates.SpectralCoord}{SpectralCoord}}
\newcommand{\astropySkyCoord}{\astropyapidoc{coordinates.SkyCoord}{SkyCoord}}
\newcommand{\astropyICRS}{\astropyapidoc{coordinates.builtin_frames.ICRS}{ICRS}}
\newcommand{\astropyGalacticLSR}{\astropyapidoc{coordinates.builtin_frames.
GalacticLSR}{GalacticLSR}}
\newcommand{\astropyAltAz}{\astropyapidoc{coordinates.builtin_frames.AltAz}{AltAz}}
\newcommand{\astropyGalactocentric}{\astropyapidoc{coordinates.builtin_frames.Galactocentric}{Galactocentric}}
\newcommand{\astropyCosmology}{\astropyapidoc{cosmology.Cosmology}{Cosmology}}
\newcommand{\astropyFlatLambdaCDM}{\astropyapidoc{cosmology.FlatLambdaCDM}{FlatLambdaCDM}}
\newcommand{\astropyFitsColumn}{\astropyapidoc{io.fits.Column}{Column}}
\newcommand{\astropyModel}{\astropyapidoc{modeling.Model}{Model}}
\newcommand{\astropyTableColumn}{\astropyapidoc{table.Column}{Column}}
\newcommand{\astropyTable}{\astropyapidoc{table.Table}{Table}}
\newcommand{\astropyQTable}{\astropyapidoc{table.QTable}{QTable}}
\newcommand{\astropyTime}{\astropyapidoc{time.Time}{Time}}
\newcommand{\astropyLeapSeconds}{\astropyapidoc{time.LeapSeconds}{LeapSeconds}}
\newcommand{\astropyDistribution}{\astropyapidoc{uncertainty.Distribution}{Distribution}}
\newcommand{\astropyUnit}{\astropyapidoc{units.Unit}{Unit}}
\newcommand{\astropyQuantity}{\astropyapidoc{units.Quantity}{Quantity}}
\newcommand{\astropyMasked}{\astropyapidoc{utils.masked.Masked}{Masked}}
\newcommand{\astropyScienceState}{\astropyapidoc{utils.state.ScienceState}{ScienceState}}

@@

% Maybe have command link to Zenodo reference instead, or include it?
\newcommand{\astropyAPE}[1]{\href{https://github.com/astropy/astropy-APEs/blob/main/APE#1.rst}{APE #1}\xspace}

@@

The \astropypkg core package is mature and stable, in that many features have
been part of the core package for years with purposefully few changes to the
software interface.  This maturity and stability allows for the broader
astronomy \python community to rely upon the \astropypkg core package and build
specialized packages within the Astropy ecosystem.  Even many non-astronomical
\python libraries have come to rely on \astropypkg. The relationship between
\astropy and the \python-using community is reciprocal, with the evolving needs
of the community and the maturation of the broader scientific \python ecosystem
driving much of the development of the \astropypkg core package.

Within \astropypkg, development may be roughly split into a few
categories: new features added, intra and inter-package interoperability,
improvements to precision, accuracy, and reproducibility, and performance
enhancements. We discuss these in turn.

Additionally, some features have been moved from the core package, either
because they are of more general use outside of \astropypkg, or because they are
too specialized and belong in an Astropy-affiliated package. An example of the
former is the copy of the IAU Standards Of Fundamental Astronomy (SOFA) software
Collection\footnote{\url{http://www.iausofa.org}} \citep{sofa} that \astropy
carried, as well as the \python wrappers for it. Since this is basic
infrastructure, with a release cycle set by SOFA, it made more sense to create
separate packages that serve as \astropypkg dependencies: \package{ERFA}
\citep{erfa} and \package{PyERFA} \citep{pyerfa}.  An example of something that
was too specialized was a Virtual Observatory sub-package; the Simple
Application Message Protocol made sense to keep (\astropysubpkg{samp}), but
other parts more properly belonged in \package{astroquery} (see
Sect.~\ref{sec:astroquery}).

\subsubsection*{New and Planned Features} \label{sec:core-features-new}

New features present a particular difficulty for a mature package such as
\astropypkg, because it is difficult to know {\em a priori} what the best
interface will be, and hard to address all possible use cases in one round of
development, even within \astropypkg\ itself. New features come in various forms,
from new submodules providing wholly new capabilities, to larger additions to
existing submodules, to large rewrites of the interface of modules. This range
is spanned by five main new features that entered the \astropypkg core since the
previous summary \citepalias{astropy:2018}. In our descriptions of each we
include how we hope the new feature will evolve. Ultimately, in a user-driven
project like \astropy, development depends primarily on the priorities of
contributors.

\paragraph{Uncertainties and Distributions} Early in our careers we were likely
all taught that measurements without units and uncertainties are meaningless.
\astropyunits provides the ability to associate numbers with units, and
propagate these correctly, using the \astropyQuantity\ class.  However, the
ability to associate and propagate uncertainties is relatively limited: in
\astropysubpkg{nddata}, there are options to associate errors with data arrays,
but these are treated as independent, i.e., covariances are not tracked.

Error propagation is difficult, and often is best approximated using Monte Carlo
methods.  The new \astropysubpkg{uncertainties} sub-package is our first step
towards enabling seamless error propagation.  The package allows one to
generate, for each variable, randomly drawn samples in a \astropyDistribution,
and then propagate these by passing them through the normal analysis, producing
a \astropyDistribution\ of final results that can be inspected. This
\astropyDistribution\ can be any type of array, including a \astropyQuantity.
The goal is to ensure \astropyDistribution\ can be seamlessly used to
instantiate other \astropypkg classes as well, such as \astropySkyCoord,
\astropyTime, and \astropyCosmology.

A plan for future development of \astropysubpkg{uncertainties} is to also
support error propagation, tracking covariances, for the case that uncertainties
are normally distributed rather than approximated by Monte-Carlo means. One
implementation problem to be solved is when to {\em stop} tracking covariances.
For instance, a simple operation such as subtracting the mean from $N$ data
points implies that all data points are now covariant with each other, i.e., one
has to carry $N\times N$ covariances.  For a large image, that becomes not just
pointless but also prohibitive in terms of memory and CPU usage.

\paragraph{Masked quantities}  It can be useful to mask bad data.  How one
supports masked data involves a number of choices, such as whether one indicates
data values are bad by a separate flag or by replacing bad data with a special
value, such as ``Not a Number'' (NaN).  In \astropypkg, both approaches have
been used: a flag for masked columns in \astropyTable, mask flags and bitmaps
for N-dimensional data in \astropysubpkg{nddata}, and replacing elements with
NaN in \astropyTime.  However masked quantities were poorly supported, making it
difficult to use masks in a \astropyQTable\ (in which \astropyQuantity\ is used
for all columns with units).

Unfortunately, the \texttt{MaskedArray} class from \package{numpy} only works
well with plain data arrays, hiding other metadata and attributes, for example
the \texttt{unit} of a \astropyQuantity. Consequently, a new \astropyMasked\
class was designed, based on a framework very similar to that of
\astropyDistribution, making it easy to create masked instances of other
classes.  In particular, \astropyMasked\ can be used to create masked quantities
for \astropyQTable, enabling I/O of masked data from different file types.

The masked quantities work largely without any changes in higher-level objects
such as \astropySkyCoord, but work is still underway to expose the mask in those
classes, as well as to use masked arrays in \astropyTime\ instead of replacing
data with NaN.

\paragraph{Time series}  From sampling a continuous variable at fixed times to
counting events binned into time windows, many different areas of astrophysics
require the manipulation of time series.  The new \astropysubpkg{timeseries}
sub-package extends the \astropyQTable\ class to tables of data as a function of
time, where the data can either represent samples or averages over particular
time bins. The new classes offer a number of special methods to manipulate time
series (folding, resampling, etc.) and to read different data formats (such as
{\em Kepler} lightcurves, etc.).

Also part of \astropysubpkg{timeseries} are common analysis routines, including
\href{https://docs.astropy.org/en/stable/timeseries/lombscargle.html}{Lomb-Scargle}
and
\href{https://docs.astropy.org/en/stable/timeseries/bls.html}{box-least-squares}
periodogrammes.

\paragraph{Spectral Coordinates} Measurements are often taken at specific
``spectral coordinates,'' be they frequencies, wavelengths, or photon energies.
\astropyQuantity\ can represent these and convert between them via dedicated
equivalencies. The new \astropySpectralCoord\ builds on \astropyQuantity\ by
providing a more straightforward interface: baking in these equivalencies and
those for Doppler velocities. Furthermore, \astropySpectralCoord\ can be made
aware of the observer and target reference frames, allowing transformation from
telescope-centric (or topocentric) frames to Barycentric or Local Standard of
Rest (LSRK and LSRD) velocity frames.

\paragraph{A High-Level Interface to World Coordinate Systems}
Astronomical data are often provided alongside information about the
correspondence between ``real-world'' and pixel coordinates. This mapping is the
essence of the ``World Coordinate System (WCS)'' concept. From its inception,
the \astropysubpkg{wcs} sub-package allowed access to WCS information provided
in, e.g., FITS files, but it became clear that other WCS standards and
representations had to be supported for new missions and observatories (e.g.,
the James Webb Space Telescope and the Rubin Observatory).  To harmonize these
needs, a new high-level interface was created based on a formal design first
proposed in an \astropy\ Proposal for Enhancement (\astropyAPE{14}). The hope is
that by having a formal design, other packages implementing WCS objects can
straightforwardly modify their classes to conform to the new interface or build
thin wrappers that conform.  The implementation in \astropysubpkg{wcs} interacts
well with other \astropypkg\ objects such as \astropySkyCoord\ and \astropyTime.

\subsubsection*{Interoperability} \label{sec:core-features-interoperability}

Because \astropy is modular and situated at the nexus between scientific
computing and astronomy, interoperability is an important focus for \astropy,
both within the various \astropypkg sub-packages and with other \python
libraries. If done right, all a person notices is that code using many different
python features and libraries ``just works.'' Here, we describe a few particular
efforts towards this goal.

\paragraph{\package{numpy} on Units}

\astropyQuantity\ is a backbone of \astropy, leveraging the power of
\package{numpy} and adding units. Previously, many \package{numpy} functions
would strip a \astropyQuantity\ of its units (or fail outright), limiting
\astropyQuantity's usefulness. Now, advances in \package{numpy} function
overloading (e.g.,
\href{https://numpy.org/neps/nep-0018-array-function-protocol.html}{NEP 18})
mean \astropyQuantity\ works with almost all \package{numpy} (v1.17+)
functions.

In the remaining gaps, the \package{numpy} and \astropyQuantity\
interoperability efforts are ongoing. For some functions, such as in
\href{https://numpy.org/doc/stable/user/basics.rec.html#module-numpy.lib.recfunctions}{\package{numpy}'s
module} for manipulating structured arrays, compatibility only requires
extending the existing \package{numpy}-\astropypkg bridge frameworks. Community
interest, in the form of a Feature Request (or better yet Pull Request) would be
sufficient to see this compatibility completed. For a few remaining functions,
discussed further in the
\href{https://docs.astropy.org/en/stable/known_issues.html#known-deficiencies}{\astropy
documentation}, the \package{numpy} framework does not yet allow for full
interoperability with \astropy. A goal of \astropy and the scientific \python
community is to enhance and implement the frameworks, allowing \astropyQuantity\
to propagate units seamlessly across the whole ecosystem of \package{numpy}-like
projects.

\paragraph{\astropypkg on Units}

Units were also integrated further within \astropypkg. For an already-defined unit-less
\astropyModel, e.g. those imported from another library, \astropymodeling can
now coerce units.
% TODO! transition words
A new module – \astropycosmologyunits – has been added to the cosmology
sub-package for defining and collecting cosmological units and equivalencies.
The unit \texttt{littleh} and equivalency
\astropyapidoc{cosmology.units.with_H0}{with\_H0} was moved from the main
\astropyunits sub-package to \astropycosmologyunits. A new unit,
\texttt{redshift}, has been added for tracking factors of cosmological redshift.
To correctly use redshift units in dimensional analysis, \texttt{redshift} is
\astropyapi{cosmology.units.dimensionless_redshift}{treated as dimensionless}
by default. To convert between redshift and other cosmological distance
measures, e.g. CMB temperature or comoving distance, the equivalency
\astropyapidoc{cosmology.units.with_redshift}{with\_redshift}
has also been added. This equivalency is actually a composite of other equivalencies,
which may be used separately.

\paragraph{Table Mixin Columns}

Within \astropypkg and \python there are numerous ways to represent and store
array-valued data. Some of the differences are historical:
\astropyapidoc{table.Column}{table.Column} and
\astropyapidoc{io.fits.Column}{io.fits.Column} are not the same. Some
differences are computational: \package{numpy}, \package{cupy}, and
\package{dask} arrays have almost identical APIs, but are optimized for
different use cases. Lastly, some differences are inherent: \astropyQuantity,
\astropyTime\, and \astropySkyCoord\ represent fundamentally different types of
objects.

We aim to make it possible for any array-valued data to be used as a column in a
table. A well defined protocol for mixin-columns has been developed for
\astropytable, allowing the original object to be used as a column with a
familiar API, and to ``round-trip'' through tables with no loss of data or
attributes. With this protocol, it is now possible to store \astropy native
objects -- including \astropyTime, \astropyQuantity, and \astropySkyCoord --
within a \astropyTable and write these to various file formats, such as FITs.
For objects not already covered by the mixin protocol, functions can be
registered with \astropyTable\ to convert any array-like object into a mixin
column. As an example, the mixin functions are used to integrate \package{dask}
arrays with \astropyTable, allowing cloud-stored or cluster-scale data to be
used as a column in a \astropyTable.

\paragraph{Astropy FITS in Time}

The FITS standard was extended to rigorously describe time coordinates in the
World Coordinate System (WCS) framework \citep{FITS-Time:2015}. Compared to
other types of coordinates in WCS, time requires more metadata: format, scale,
position, reference, etc. This metadata had to be manually specified and the
nuances understood by the user. \astropyfits could read this data as a standard
\astropyFitsColumn\, but would not interpret the time-related metadata and
attributes. Through the support of the
\href{https://summerofcode.withgoogle.com/archive/2017/projects/4778482366152704}{Google
Summer of Code} 2017 program\footnote{ To learn more about this project, please
see the final report,
\href{https://aaryapatil.wordpress.com/2017/08/28/a-mixin-protocol-for-seamless-interoperability/}{A
mixin protocol for seamless interoperability}}, the \astropyfits package now
interprets time data correctly, using \astropyTime\ as a `mixin' column.  For
backward compatibility with manual systems, this feature may be turned off.

\paragraph{Persistent Storage}

\astropy's efforts to increase support for column types in a table is mirrored
by efforts to expand storage format options. New formats have been added, and
existing formats updated to support more column types.

\astropy now supports reading and writing tables in the American Astronomical
Society Journals' Machine-Readable Table (MRT) format. This ASCII format has
long been missing and as added in this year's Google Summer of Code project adds
\href{https://docs.astropy.org/en/stable/io/ascii/write.html#cds-mrt-format}{MRT}
to \astropyTable\ I/O. In addition, \astropyTable\ may also read from and write
to \href{https://asdf-standard.readthedocs.io/en/latest/}{ASDF},
\href{https://docs.astropy.org/en/stable/io/unified.html#table-io-parquet}{Parquet},
and \href{https://wwwastro.msfc.nasa.gov/qdp/}{QDP} formats.

The \href{https://github.com/astropy/astropy-APEs/blob/main/APE6.rst}{ECSV
standard} has been updated to version 1.0, adding support three additional data
subtypes. First, it adds multidimensional column data (both masked and unmasked)
with fixed dimensions in all table cells. Second, it adds multidimensional
column data with variable-dimension arrays similar to FITS variable-length
arrays. Third, it adds object-type columns with simple \python objects.

As mentioned previously, \package{dask} arrays may be used as a mixin column in
\astropyTable. Now, \package{dask} may also be the data array in
\astropyapidoc{io.fits.PrimaryHDU}{FITS HDU} and if written to disk, the array
will be computed while written, avoiding excessive memory use. \astropyTable\ can
now be appended to an existing FITS file, and \package{dask} mixin columns
interoperate seamlessly between the two table types.

\paragraph{Unified I/O architecture}

% TODO! better transition
\astropyiosubpkg{registry} is a powerful way to define input and output (I/O)
functions for \astropypkg objects, such as reading from or writing to a file,
following a given standard. \astropy uses this internally for the I/O methods in
\astropyTable\ and \astropyCosmology, and users can register custom I/O to
extend the options on these classes.

As of Astropy 5.0, the I/O registry submodule has been generalized to enable a
number of new use cases. for instance, the class-based architecture allows for
the creation of custom registries. One application is in \astropycosmology,
which has two different \textit{kinds} of registries for \astropyCosmology: one
for reading and writing files, and another for converting between python
objects. Using \astropyiosubpkg{registry} for custom classes means a developer
does not have to create a whole new I/O system and can offer to the user a
unified and familiar API.

\subsection*{Precision, Accuracy, \& Reproducibility}
\label{sec:core-features-precision_accuracy_repreducibility}

An important focus of \astropy has been to increase the precision and accuracy
of its functionality, while, where appropriate, making sure results obtained
using older versions of \astropypkg are reproducible. Within each section below
we include in our descriptions how old results may be reproduced. The primary
means are with options in configuration files and settings on runtime configuration
objects called \astropyScienceState. Notable improvements to \astropypkg have
been made to \astropytime, \astropyconstants, \astropycoordinates, and
\astropycosmology, and we describe each in turn.

\paragraph{Time}
In astronomy, time accuracy down to the (nano)second is frequently important.
Whether for planning observations, crunching pulsar or VLBI data, missing
seconds meaningfully impact results. Before, when using \astropyTime\ in
\astropytime leap seconds had to be manually applied; now, leap seconds are
applied automatically. Moreover, \astropypkg updates internal time data-files to
ensure that the correct leap second adjustment is always used. For
reproducibility, e.g. with an old code, \astropyLeapSeconds\ may also be
manually applied.

\paragraph{Constants}

Measurements of physical constants (and the units defined from them) improve
over time. Periodically the standardized systems of units and constants are
updated to reflect these improvements. For instance, in 2019 the SI system was
redefined \citep{NIST2019}, with an accompanying update to the physical
constants in
\href{https://codata.org/initiatives/data-science-and-stewardship/fundamental-physical-constants/}{\texttt{SI/CODATA
2018}}. \astropyconstants now defaults to use the \texttt{SI/CODATA 2018}
values, with the units in \astropyunits based on these constants. Most of
\astropypkg and affiliate packages build upon \astropyunits, so this update
effects the entire \astropy ecosystem.

For reproducibility, \astropypkg allows the constants' (and therefore units')
definitions to be rolled back to prior values, e.g. to the \texttt{SI/CODATA
2014} values. For work sensitive to the values of the fundamental constants we
recommend including with the work an \astropy configuration file, which
specifies the set of constants used.

\paragraph{Cosmology}
% TODO! At least one good sentence about "what is the benefit of
% astropy.cosmology"

\astropycosmology contains classes for representing cosmological models. Bundled
with the classes are commonly used \astropyCosmology\ realizations, e.g.,
best-fit measurements from
\href{https://lambda.gsfc.nasa.gov/product/map/current/}{WMAP}
\citep{WMAP2003} and
\href{https://www.nasa.gov/mission_pages/planck}{Planck}
\citep{PlanckMission:2006}.

When these important missions publish new measurements, \astropycosmology is
updated to include these results as realizations of the appropriate class
(generally \astropyFlatLambdaCDM). The models used by WMAP and Planck do not
always exactly correspond to \astropycosmology classes. For instance, the Planck
2018 results \citep{Planck2018VI:2020} include massive neutrinos in
$\Omega_{matter,0}$, while in \astropyFlatLambdaCDM\ this mass contribution is
stored in a separate parameter. Consequently, while some parameter values appear
different from the source paper, the \astropyCosmology\ realizations correctly
reproduce the cosmological models.

Since \citealt{astropy:2018} the following cosmology realizations have been
added:
\href{http://docs.astropy.org/en/stable/api/astropy.cosmology.WMAP1.html}{WMAP First Year}
\citep{WMAP1Year:2003};
\href{http://docs.astropy.org/en/stable/api/astropy.cosmology.WMAP3.html}{WMAP
Three Year} \citep{WMAP3Year:2007};
\href{http://docs.astropy.org/en/stable/api/astropy.cosmology.Planck15.html}{Planck 2015}
\citep{Planck2015XIII:2016}; and the
\href{http://docs.astropy.org/en/stable/api/astropy.cosmology.Planck18.html}{Planck
2018} best-fit cosmological parameters \citep{Planck2018VI:2020}.

Astropy provides a
\href{http://docs.astropy.org/en/stable/api/astropy.cosmology.default_cosmology.html}{configurable
default cosmology}, which is used in calculations done in a cosmological
context. The default \astropyCosmology\ has been updated to the Planck 2018
parameters. For reproducibility, this cosmology may be set to old defaults, such
as the WMAP Three Year values. The default cosmology is dynamically configurable
using \astropyScienceState, meaning a set of calculations may be run using
different assumed cosmologies, and results compared between the two.

\paragraph{Coordinates}

Being able to describe the position of objects, i.e., their coordinates, is
fundamental across many knowledge domains. \astropycoordinates allows one to
work with low-level positional and velocity data all the way to high-level
coordinate objects with reference frames, atmospheric information, and more. A
central feature of \astropycoordinates is the ability to transform data between
reference frames, e.g. \astropyICRS\ \citep{ICRS:1997} to \astropyGalacticLSR\
\citep{GalacticLSR:2010}. However, there have been some historical limitations
for spatially proximate transformations, impacting the usefulness of
\astropycoordinates for ground-based telescopes and astronometry within the
solar system. Transformations in the \astropyAltAz\ frame were reasonably
precise for very distant objects, but wrong by up to several arcseconds for
e.g., the location of the moon. Now these transformations are much more precise,
down to the milliarcsecond level. Similar precision improvements were made to
the
\astropyapi{coordinates.builtin_frames.HADec}{Hour Angle-Declination}
frame transformations. Additionally,
\astropyapi{coordinates.builtin_frames.BaseEclipticFrame}{Ecliptic frames}
and associated transformations have been updated to correctly reflect the “true”
and “mean” terminology.

For Galactic astronomers, the \astropyGalactocentric\
\astropyapi{coordinates.galactocentric_frame_defaults}{frame defaults}
have been updated to include more recent measurements (tabulated in the
\href{https://docs.astropy.org/en/stable/api/astropy.coordinates.galactocentric_frame_defaults.html#astropy.coordinates.galactocentric_frame_defaults.references}{\texttt{frame\_defaults.references}}
attribute). For reproducibility, old definitions are still available.

\subsection*{Performance} \label{sec:core-features-performance}

As features are added and corner cases dealt with, performance often suffers.
Hence, pull requests that improve performance are welcomed, and were even the
main goal for one release cycle.  For most sub-packages, performance was
improved in the \python\ code, but for some, the most time-critical pieces were
rewritten in~C (e.g., for convolution of images, sigma-clipping, and converting
time strings to binary).  Furthermore, a particular effort was made to ensure
\astropypkg\ is thread-safe, so that it can be used on supercomputer clusters.

>>>>

% Links to documentation: subpackages
\newcommand{\astropysubpkg}[1]{\href{http://docs.astropy.org/en/stable/#1/index.html}{\texttt{astropy.#1}}\xspace}
\newcommand{\astropyiosubpkg}[1]{\href{http://docs.astropy.org/en/stable/io/#1/index.html}{\texttt{astropy.io.#1}}\xspace}
\newcommand{\astropywcsaxes}{\href{http://docs.astropy.org/en/stable/visualization/wcsaxes/index.html}{\texttt{astropy.visualization.wcsaxes}}\xspace}
\newcommand{\astropycosmology}{\astropysubpkg{cosmology}}
\newcommand{\astropycosmologyunits}{\href{https://docs.astropy.org/en/stable/cosmology/units.html}{\texttt{astropy.cosmology.units}}}
\newcommand{\astropyunits}{\astropysubpkg{units}}
\newcommand{\astropycoordinates}{\astropysubpkg{coordinates}}
\newcommand{\astropyconstants}{\astropysubpkg{constants}}
\newcommand{\astropytime}{\astropysubpkg{time}}
\newcommand{\astropymodeling}{\astropysubpkg{modeling}}

@@

\newcommand{\astropySpectralCoord}{\astropyapidoc{coordinates.SpectralCoord}{SpectralCoord}}
\newcommand{\astropySkyCoord}{\astropyapidoc{coordinates.SkyCoord}{SkyCoord}}
\newcommand{\astropyICRS}{\astropyapidoc{coordinates.builtin_frames.ICRS}{ICRS}}
\newcommand{\astropyGalacticLSR}{\astropyapidoc{coordinates.builtin_frames.
GalacticLSR}{GalacticLSR}}
\newcommand{\astropyAltAz}{\astropyapidoc{coordinates.builtin_frames.AltAz}{AltAz}}
\newcommand{\astropyGalactocentric}{\astropyapidoc{coordinates.builtin_frames.Galactocentric}{Galactocentric}}
\newcommand{\astropyCosmology}{\astropyapidoc{cosmology.Cosmology}{Cosmology}}
\newcommand{\astropyFlatLambdaCDM}{\astropyapidoc{cosmology.FlatLambdaCDM}{FlatLambdaCDM}}
\newcommand{\astropyTable}{\astropyapidoc{table.Table}{Table}}
\newcommand{\astropyQTable}{\astropyapidoc{table.QTable}{QTable}}
\newcommand{\astropyTime}{\astropyapidoc{time.Time}{Time}}
\newcommand{\astropyLeapSeconds}{\astropyapidoc{time.LeapSeconds}{LeapSeconds}}
\newcommand{\astropyDistribution}{\astropyapidoc{uncertainty.Distribution}{Distribution}}
\newcommand{\astropyUnit}{\astropyapidoc{units.Unit}{Unit}}
\newcommand{\astropyQuantity}{\astropyapidoc{units.Quantity}{Quantity}}
\newcommand{\astropyScienceState}{\astropyapidoc{utils.state.ScienceState}{ScienceState}}

@@

@@

Summarize major pieces of development since the last paper (since v2.0). A few
ideas (taken from What's New pages) below, but feel free to expand on or remove
from this list:
\begin{itemize}
    \item Support for representing and transforming velocity data in coordinates, epoch propagation (v3.0)
    \item Improved support for astronomical time series: TimeSeries object (v3.2), Box Least Squares periodogram (v3.1)
    \item Overall improved support of Quantity throughout numpy (v4.0) and scipy
    \item Native support for Time, Quantity, and SkyCoord objects in Astropy tables
    \item APE 14 WCS
    \item A new SpectralCoord class for representing and transforming spectral quantities
    \item Support for dask/large arrays in several places in the core package (4.1, 5.0, and maybe other versions)
\end{itemize}
